---
title: "The Lasting Shadows of Economic Downturns"
subtitle: "Unveiling the 1980–1982 Recession's Long-Run Effects on Unemployment, Personal Income and Business Pattern in America"
author: "Heyucheng Zhang, Jiwon Choi"
thanks: "Code and data are available at: https://github.com/jwonc4602/Long-Run-Impacts-of-Recession"
date: "12 February 2024"
date-format: "D MMMM YYYY"
abstract: "This study examines the long-term impacts of the 1980–1982 recession, focusing on changes in unemployment rates and personal income in America to assess how economic downturns influence decisions regarding business patterns and human capital investment. Our analysis of historical data reveals that the recession had enduring negative effects on employment and income levels, suggesting that such economic challenges may lead individuals to another business trends as a means to improve their marketability and financial stability. These findings emphasizes the importance of incorporating long-term considerations into economic policy-making, highlighting the need for strategies that mitigate the adverse effects of recessions while fostering investment in human capital."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false
#| warning: false
#### Workspace set up ####
# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("here")
# install.packages("janitor")
# install.packages("kableExtra")

library(tidyverse) # data-related packages
library(knitr) # make tables
library(here) # manage file paths
library(janitor) # clean data
library(kableExtra) # table-related packages

#### Read in cleaned data ####
cleaned_unemployment <- read_csv(file = here("outputs/data/cleaned_unemployment.csv"), show_col_types = FALSE)
cleaned_four_states <- read_csv(file = here("outputs/data/cleaned_four_states.csv"), show_col_types = FALSE)
cleaned_personal_income <- read_csv(file = here("outputs/data/cleaned_personal_income.csv"), show_col_types = FALSE)
cleaned_bussiness_pattern1 <- read_csv(file = here("outputs/data/cleaned_bussiness_pattern1.csv"), show_col_types = FALSE)
cleaned_bussiness_pattern2 <- read_csv(file = here("outputs/data/cleaned_bussiness_pattern2.csv"), show_col_types = FALSE)
```

# Introduction

*This reproduction was performed after a replication on the Social Science Reproduction platform: [**link here**](update%20the%20link!!!)*

# Data
This paper replicates and extends the analyses from the original paper, 'The Long-Run Effects of Recessions on Education and Income' [@10.1257/app.20180055] by using a range of datasets obtained from various publicly accessible sources. Stuart has categorized the datasets into distinct files, reflecting their respective roles in the analysis.

## Sources
We used data from two sources for our research. First, we utilized data from the Bureau of Economic Analysis (BEA) Regional Economic Accounts[@BEA]. These datasets were downloaded in March 2015 from <https://apps.bea.gov/regional/downloadzip.cfm> by Stuart. The BEA occasionally revises the methodology used to generate these data, and accordingly, updates all files to reflect these changes. As a result, data downloaded on subsequent dates might present some variations. The datasets of particular interest included 'CA1: Personal Income Summary: Personal Income, Population, Per Capita Personal Income', which covers Personal Income, Population, and Per Capita Personal Income, and 'CA25: Total Full-Time and Part-Time Employment by SIC Industry', offering detailed insights into income components and industry-specific earnings.

Second, we obtained unemployment data for the years 1976-1989 which Stuart obtained through email communication with Brian Hannon of the BLS from the Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics [@BLS]. This data, covering various counties, is compiled in the file 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' and is available through the Inter-university Consortium for Political and Social Research.

The datasets 'CA1_1969_2013_ALL.csv' and 'CA25_1969_2000_ALL.csv' from the BEA, along with the 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' from the BLS are crucial for analyzing personal income, business patterns, and unemployment trends in our study.

## Methodology and Measurement
The data collection methodology is designed to analyze unemployment, personal income, and business trends comprehensively. Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like `tidyverse` [@tidyverse], `ggplot2` [@ggplot], `dplyr` [@dplyr], `readxl` [@readxl], `tibble` [@tibble], `janitor` [@janitor], `knitr` [@knitr], and `here` [@here] for support. The data is mainly numeric, unemployment rate, personal income per capitia, or counting the number of each business patterns.

Initial data cleaning from 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' focused on nationwide unemployment rates (see @tbl-first-cleaned-unemployment-data), but due to the breadth of data, the study narrowed its focus to the four most populous states in the USA: California, Florida, New York, and Texas (see @tbl-cleaned-unemployment-data-focusing-four-states).

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-first-cleaned-unemployment-data
#| tbl-cap: Sample of First Cleaned Unemployment Data
kable(cleaned_four_states, format = "latex", booktabs = TRUE, align = "c") |>
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), font_size = 7, full_width = FALSE)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-cleaned-unemployment-data-focusing-four-states
#| tbl-cap: Cleaned Unemployment Data Focusing Four Most Populated States
kable(head(cleaned_unemployment, 30), format = "latex", booktabs = TRUE, align = "c") |>
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
```
We also utilized the dataset 'CA1_1969_2013_ALL.csv' to extract three specific rows detailing per capita personal income (in dollars) from 1969 to 2013. To ensure consistency with unemployment rate data, we narrowed the focus to the years 1979 to 1989 (see @tbl-personal-income).
```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-personal-income
#| tbl-cap: Cleaned Personal Income Per Capita in Dollars
kable(cleaned_personal_income, format = "latex", booktabs = TRUE, align = "c") |>
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), font_size = 7, full_width = FALSE)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-4
#| tbl-cap: Sample 4
kable(head(cleaned_bussiness_pattern1, 10), format = "latex", booktabs = TRUE, align = "c") |>
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-5
#| tbl-cap: Sample 5
kable(head(cleaned_bussiness_pattern2, 10), format = "latex", booktabs = TRUE, align = "c") |>
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
```

# Results

```{r fig.width=10, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| label: fig-1
#| fig-cap: plot 1
### the line graphs of average unemployment rate by four most populated states###
plot_data1 <-
  cleaned_four_states  |>
  select(year, state_name, average_unemployment_rate) |>
  arrange(year, state_name)

# Define custom colors for each state
state_colors <- c("California" = "red", "Florida" = "green", "New York" = "blue", "Texas" = "yellow")

# Create a line plot with custom styles
plot1 <- ggplot(plot_data1, aes(x = year, y = average_unemployment_rate, color = state_name)) +
  geom_line(size = 1) +
  scale_color_manual(values = state_colors) + # Set custom colors for lines
  geom_vline(xintercept = 1980, linetype="dashed", color = "red", size = 1) +
  geom_vline(xintercept = 1982, linetype="dashed", color = "red", size = 1) +
  annotate("text", x = 1980, y = max(plot_data1$average_unemployment_rate), label = "Recession Start", vjust = 2) +
  annotate("text", x = 1982, y = max(plot_data1$average_unemployment_rate), label = "Recession End", vjust = -0.5) +
  labs(title = "Average Unemployment Rate by Four Most Populated States and Year",
       x = "Year", y = "Average Unemployment Rate") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels (optional)
        legend.position = "bottom") # Move legend to the bottom

# Display the plot
print(plot1)

```

```{r fig.width=10, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| label: fig-2
#| fig-cap: plot 2
### the bar graph change of unemployment for four major states ###
### something like simulated plot2 ###
plot_data2 <-
  cleaned_four_states  |>
  select(year, state_name, average_unemployment_rate) |>
  arrange(year, state_name) |>
  filter(year %in% c(1976, 1982, 1989))

# Define custom colors for the years
year_colors <- c("1976" = "green", "1982" = "red", "1989" = "blue")

# Create a bar plot with custom styles
plot2 <- ggplot(plot_data2, aes(x = state_name, y = average_unemployment_rate, fill = as.factor(year))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = year_colors) +  # Set custom colors for bars
  labs(title = "Change of Unemployment for Four Major States",
       x = "State", y = "Change in Unemployment Rate") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the plot title
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels (optional)
    legend.position = "bottom"  # Move legend to the bottom
  )

# Display the plot
print(plot2)
```

```{r fig.width=10, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| label: fig-3
#| fig-cap: plot 3
### the line graph of personal income per capita by year ###
plot_data3 <-
  cleaned_personal_income  |>
  filter(year %in% 1976:1989)

# Create a line plot with custom styles
plot3 <- ggplot(plot_data3, aes(x = year, y = `Per capita personal income (dollars) 2/`)) +
  geom_line(color = "forestgreen", size = 1) +  # Set line color and size
  labs(x = "Year", y = "Income Per Capita", 
       title = "Personal Income Per Capita by Year") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the plot title
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels (optional)
  )

# Display the plot
print(plot3)
```

```{r fig.width=10, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| label: fig-4
#| fig-cap: plot 4
plot_data4 <- cleaned_bussiness_pattern1 |>
  pivot_longer(
    cols = -year,
    names_to = "EmploymentType",
    values_to = "Count"
  ) |>
  filter(year %in% 1976:1989)

# Create a line plot
plot4 <- ggplot(plot_data4, aes(x = year, y = Count, color = EmploymentType)) +
  geom_line() +
  labs(
    title = "Business Pattern 1",
    x = "Year",
    y = "Count",
    color = "Employment Type"
  ) + theme_minimal()

# Display the plot
print(plot4)
```

```{r fig.width=10, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| label: fig-5
#| fig-cap: plot 5
plot_data5 <- cleaned_bussiness_pattern2 |>
  pivot_longer(
    cols = -year,
    names_to = "EmploymentType",
    values_to = "Count"
  ) |>
  filter(year %in% 1976:1989)

# Create a line plot
plot5 <- ggplot(plot_data5, aes(x = year, y = Count, color = EmploymentType)) +
  geom_line() +
  labs(
    title = "Business Pattern 1",
    x = "Year",
    y = "Count",
    color = "Employment Type"
  ) + theme_minimal()

# Display the plot
print(plot5)
```

# Discussion

## How Recession Impacts on Unemployment

## How Recession Impacts on Personal Income

## How Recession Changes the Business Patterns

## Limitations and Next Steps
Due to limitations in accessing relevant datasets, this paper will not explore the diverse social impacts of recessions, specifically on birth rates and education, which were initially of primary interest. Drawing from Stuart's foundational research, we quantitatively assess the prolonged impacts of the 1980–1982 recession on individuals who were in their formative years—children, adolescents, and young adults—at the onset of this economic downturn. Stuart's findings reveal significant long-term detriments to educational attainment and income levels. Specifically, for individuals aged 0–10 in 1979, a 10 percent reduction in per capita earnings from 1979 to 1982 within their birth county correlates with a 4.7 percentage point (14.6 percent) reduction in the attainment of four-year college degrees, a decrease in earned income by $2,100 (5.2 percent), and a 1.6 percentage point (13.0 percent) increase in the likelihood of living in poverty during 2000–2013. The adverse effects on college graduation rates are notably pronounced and consistently significant for individuals aged 0–13 in 1979, and diminish to small and statistically non-significant levels for those aged 14–22. This pattern indicates that the recession's impact likely stems from a reduction in childhood human capital or a sustained decrease in parental financial capacity to support higher education. Future research will address the effects on education and birth rates once relevant and accessible datasets become available.

\newpage

# References
